\chapter[Curriculum Vit\ae]{
\hrule
\medskip\medskip\medskip
Curriculum Vit\ae
\medskip\medskip\medskip
\hrule
}

\section{Contact Information}

40, Rue Colin\\
Villeurbanne, 69100, France.\\
Mobile Number: +33-(0)-638305900\\
e-mail: romulo.pinho@lyon.unicancer.fr\\
URL: www.creatis.insa-lyon.fr/rio/RomuloPinho

\section{Biographical Data}

Birth date: March, 3rd, 1977\\
Place of Birth: Petr\'opolis, RJ, Brazil\\
Citizenship: Brazilian\\
Marital status: Married

\section{Current Position}

\begin{longtable}{p{0.25\textwidth} p{0.6\textwidth}}
January, 2011 \newline present & Post-doc Research Engineer \newline
Centre L\'eon B\'erard, Lyon, France \newline
Design, implementation, and deployment of a medical imaging system for image guided
radiation therapy against lung cancer. The aim is to carry out a clinical trial for the use of new
image registration techniques and their influence on radiation therapy against moving tumours. 
Multi-platform development using C++, CMake, ITK, VTK, QT, Python, and shell scripts.\\
\\
\end{longtable}


\section{Education}

\begin{longtable}{l p{0.6\columnwidth}}
November, 2010 & Ph.D. in Physics. \newline 
Universiteit Antwerpen, Antwerp, Belgium \newline 
Specialization: Medical Image Processing \newline 
Thesis: A Decision Support System for the Assessment and Stenting of Tracheal Stenosis. \\
\\
May, 2001 & M.Sc. in Software Engineering. \newline 
Universidade Federal do Rio de Janeiro, Rio de Janeiro, Brazil \newline 
Specialization: Computer Graphics \newline 
Thesis: Three Dimensional Reconstruction of Head Models from Magnetic Resonance Images. \\
\\
September, 1998 & B.Sc.  in Computer Science \newline 
Universidade Federal Fluminense, Niter\'oi, Brazil \newline 
Specialization: Computer Science.\\
\end{longtable}


\section{Honours, Awards and Certificates}

\begin{tabular}{p{0.25\columnwidth} p{0.6\columnwidth}}
Las Vegas, April, 2005 & ScrumMaster Certificate.
\end{tabular}

\section{Languages}

\begin{tabular}{l p{0.6\columnwidth}}
Portuguese & native\\
English & fluent (non-native)\\
French & intermediate level\\
Dutch & basic level\\
\end{tabular}

\input{Contributions}

\section{Academic Experience}

\subsection{Teaching}
\begin{tabular}{p{0.25\columnwidth} p{0.6\columnwidth}}
September, 2011 \newline June, 2012 & Teacher assistant \newline
Subject: Operating Systems \newline
(in French, 16h, practice sessions) \newline
Institution: INSA-Lyon\\
\\
September, 2011 \newline June, 2012 & Teacher assistant \newline
Subject: Object Oriented Programming \newline
(in English, 45h, theory and practice sessions) \newline
Institution: INSA-Lyon\\
\\
\end{tabular}

\subsection{Supervision of Theses}
\begin{tabular}{p{0.25\columnwidth} p{0.6\columnwidth}}
January, 2009 \newline June, 2009 & B.Sc. Thesis of Stijn Peeters (co-supervision) \newline
Title: Automatic segmentation of the airways: Morphological study of the trachea (in -Dutch) \newline
Institution: University of Antwerp\\
\\
January, 2009 \newline June, 2009 & B.Sc. Thesis of Sten Luyckx (co-supervision) \newline
Title: Automatic segmentation of the airways in 3D CT using cylindrical shape modelling (in Dutch) \newline
Institution: University of Antwerp\\
\end{tabular}

\subsection{Research Projects}
\begin{tabular}{p{0.25\columnwidth} p{0.65\columnwidth}}
April, 2010 \newline December, 2010 & CIMI -- IBBT \newline 
Responsible for and lead researcher of Task 2.1 of WP 3 \newline
Development of cache and pre-fetching techniques for out-of-core processing and visualization of large microscopic images. \\
\\
April, 2010 \newline December, 2010 & Segmentor -- IBBT \newline
Researcher \newline
Development of semi-automatic segmentation techniques based on haptics and deformable models for the segmentation of tracheal stenosis.
\end{tabular}


\section{Professional Experience}

\begin{longtable}{p{0.25\columnwidth} p{0.65\columnwidth}}

January, 2001 \newline April, 2006 & TV Systems Software Engineer/Researcher \newline
R\&D Department, TV Globo Ltd. \newline
Design and development of software for general acquisition,
compression, storage, processing, and transmission of
video and audio signals. Applications were implemented in C/C++ for Win32 and
Linux platforms, using various SDKs and libraries
including DirectShow SDK, Windows Media SDK, Matrox
DigiSuite SDK, MFC, Video4Linux, MySQL.\\
\\
October, 2001 \newline May, 2003 & Software Engineer/Researcher\newline
LAMEC -- Laboratory of Computational Mechanics,\newline
Federal University of Rio de Janeiro \newline
Design and development of a C++, OpenGL/QT-based scientific
visualization and geometric modelling system to build 3d 
models out of simple 3d primitives, to be used in numerical computations.\\
\\
August, 2000 \newline December, 2000 & Software Engineer\newline
Publintel S. A.\newline
Development and maintenance of GIS applications. Translation
of existing GUI code to the Win32 environment. Design and
implementation of a logistics database for automatic vehicle 
localization, using SQL-Server, C++, STL, MFC, and ADO.\\
\\
March, 2000 \newline August, 2000 & Technology Researcher/Programmer\newline
Montreal Inform\'atica\newline
Design and implementation of Speech Recognition and
biometrics (fingerprint identification) systems. These 
applications were implemented in C++ and Delphi, using 
the ORACLE DBMS.\\
\\
January, 2000 \newline March, 2000 & System Analyst/Programmer\newline
Delta de Friburgo\newline
I took part in the design of a profit optimization
system for the Brazilian Petroleum Company
(PETROBRAS), using C++ and the ORACLE DBMS.\\
\\
April, 1999 \newline January, 2000 & Programmer\newline
Brainstorming Ass. e Plan. em Inform\'atica\newline
Lead programmer of a database system to control 
the use of the necessary material for the repair of ships, weapons 
and for general services inside a Brazilian Navy base in Rio de Janeiro.
Development in Delphi/ORACLE.\\
\\
January, 1998 \newline January, 1999 & Game Programmer\newline
Z-Movie Studio.\newline
Development of 3D educational computer games. Design and
implementation of 3D engine, character animation, GUIs and game logic.
Implementations in C++ for DirectX 6.0.\\
\\
October, 1996 \newline January, 1998 & Programmer\newline
ADD-Labs -- Universidade Federal Fluminense (Laboratory of Active Document Design)\newline
Development of a prototype Virtual Reality system built in C/C++ to the Brazilian Petroleum 
Company (PETROBRAS). The objective was to provide the user with a 3D 
visualization of oil extraction fields.\\
\end{longtable}

